Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-19 analysis/paper/paper.Rmd
more texts and one plot

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005]. Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale. Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems (Dietler 1997; Junker 1993; Mitchell 2000; Silliman 2005). Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices (Given 2004; Mullins 2011; Silliman 2001; Rubertone 2000; Torrence 2000; Torrence and Clarke 2000; Voss 2005).

@@

ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency",
       x = "Categories") +
  theme_minimal(base_size = 14)
```

@@

## Frequency for spatial pattern
Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period. It shows that ornaments concentrate on one specific square in the pre-European contact, and more ubiquitous during the European contact and

 - for each time period, spatial patterning

@@

```{r plot-spatial-distribution}
# filter golden bead
ornaments_period_join_gold <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Golden bead')

@@

# plot by period
ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Spatial Distribution of Agate Beads",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))
```

>>>>

Body ornaments are usually viewed as materials that convey social meaning or representations in a culture through the body as the scene to display [@Joyce2005]. Foreign ornaments including glass beads, agate or carnelian beads, and metal objects first appeared in Taiwan around 1,800 years ago through the regional exchange network in Southeast Asia [@WangandLiu2007]. Those new types of ornaments gradually replaced the earlier jade ornamental system in Taiwan since the transition of Late Neolithic and the Iron Age. We observed a greater quantity and variety of foreign goods in the early 17th century during the European canonization period that involved the whole island in an international trade network on a large scale. Foreign trade goods introduced into local indigenous societies in a colonial or imperial contexts usually had an impact on local societies in many parts of the world that might cause transformations of indigenous economic, cultural, and socio-political systems [@Dietler1997;@Junker1993;@Mitchell2000;@Silliman2005). Contrary to as passive receptors of imperial power, the agency of indigenous people in the colonial contexts has been broadly discussed that emphasize the ability of active adoptions of foreign materials, negotiation between colonized and colonizer, and resistance of indigenous people through daily cultural practices [@Given2004;@Mullins2011;@Silliman2001;@Rubertone2000;@Torrence2000;@TorrenceandClarke2000;@Voss2005).

@@

ggplot(ornaments_count) +
  geom_col(aes(Categories, n)) +
  facet_wrap(~ period) +
  coord_flip() +
  labs(y = "Frequency", x = "Categories") +
  theme_minimal(base_size = 14)

# stack histogram
ggplot(ornaments_count) +
  geom_col(aes(Categories, n, fill = period)) +
  labs(y = "Frequency", x = "Categories") +
  coord_flip() +
  theme_minimal(base_size = 12) +
  scale_fill_viridis_d()
```

@@

## Frequency for spatial pattern
Figure \@ref(fig:plot-all-spatial-pattern) presents the spatial distribution of the ornaments for each time period. It shows that one square unearthed more ornaments than others in the pre-European contact. During the European contact period, the distribution is more ubiquitous with some concentration on northern part, and more even during Chinese contact period. Further look at the distribution for each major category individually, it shows that some squares have higher number of ornaments during the European contact period, including golden bead, agate bead, metal ring, and bell. However, there is no consistent pattern of concentration for different ornaments across the area. The squares with high number of ornaments distributed separately and independently without clear correlation. For example, more golden beads were found at the northern part, while more agate beads concentrated in the middle part.

Golden beads and

@@

```{r plot-spatial-pattern-each}
# filter golden bead
ornaments_period_join_gold <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Golden bead')

@@

# plot by period
ggplot(ornaments_period_join_gold) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_gold, aes(fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Spatial Distribution of Agate Beads",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

# filter metal ring
ornaments_period_join_metal <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Metal ring')

# plot by period
ggplot(ornaments_period_join_metal) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_metal, aes(fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Spatial Distribution of Metal Ring",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))

# filter metal ring
ornaments_period_join_bell <-
  ornaments_period_join %>% 
  filter(Categories %in% 'Bell')

# plot by period
ggplot(ornaments_period_join_bell) +
  geom_sf(data = sample_units, fill = NA) +
  geom_sf(data = ornaments_period_join_bell, aes(fill = n), alpha = 0.9) +
  facet_wrap(~ period) +
  labs(title = "Spatial Distribution of Copper Bell",
       fill = "frequency") +
  scale_fill_gradient(high = "black", low = "grey") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11))
```
